Sweat electrolytes and CFTR activity in the era of CFTR modulators

### 2022-09-07 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd

```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Edith Zemanick/Sweat Test"
}
knitr::opts_knit$set(root.dir = home_dir)
rm(home_dir)
source("~/GitHub/CF/get_genotype_class_severity.R")
```

```{r}
# Import data
df = read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv",na.strings = "")
df$Diagnosis = factor(df$Diagnosis,
                      levels = c("CFTR-related disorder","CRMS","Cystic Fibrosis","Non-CF","Normal"),
                      labels = c("CFTR-related disorder","CRMS","Cystic Fibrosis","Non-CF","Non-CF"))
# Filter per Maxie's code
df = df %>% 
  filter(Age.at.Test.in.Years < 50,
         !(Diagnosis == "Non-CF" & PatientID %in% 
             c(22,23,24,29,31,474,483,484,485,505,510,785,791,792,793))) %>% 
  distinct(.) %>% filter(Sex != "",Demographics.Available == "TRUE") %>%
  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks*7)
# Add modulator info
cftr = read.csv("./Data_Raw/Copy of Report CFTR Modulators.csv")
df$Modulator = apply(df,1,function(r){
  mod = cftr[cftr$Patient.ID == as.numeric(r["PatientID"]),]
  if (any(duplicated(mod$Age.at.Start..years.))){
    mod = mod[-which(duplicated(mod$Age.at.Start..years.)),]
  }
  mod = mod[order(mod$Age.at.Start..years.),]
  age = as.numeric(r["Age.at.Test.in.Years"])
  as.character(cut(age,c(-Inf,mod$Age.at.Start..years.,Inf),
                   labels = c("None",mod$Modulator),right = F))
})
df = left_join(df,cftr,by = c("PatientID" = "Patient.ID"))
# Genotype info
df$geno1 = get_cf_genotype_class_severity(df$Genotypes1)
df$geno2 = get_cf_genotype_class_severity(df$Genotypes2)
df$geno1[df$Diagnosis == "Non-CF"] = "Not CF"
df$geno2[df$Diagnosis == "Non-CF"] = "Not CF"
df$geno_risk = paste(df$geno1,df$geno2,sep = "_")
df$geno_risk = factor(df$geno_risk,
                      levels = c("Mild_Severe","Mild_Unknown","Not CF_Not CF","Severe_Mild","Mild_Mild",
                                 "Severe_Severe","Severe_Unknown","Unknown_Severe","Unknown_Unknown"),
                      labels = c("Low","Low","Not CF","Low","Low","High","Unknown","Unknown","Unknown"))
```

# Table 1: Patient Demographics and Sample Characteristics for sweat tests done on individuals not on CFTR modulators

```{r results='asis'}
t1_df = df %>% filter(Age.at.Test.in.Days >= 2) %>%
  group_by(PatientID) %>% arrange(Age.at.Test.in.Weeks) %>%  filter(row_number()==1)
t1 = tableby(Method ~ Sex+Age.at.Test.in.Years+geno_risk,data = t1_df)
summary(t1)
```

>>>>

```{r setup, include=FALSE}
library(readxl)
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Documents/Work/CF/Edith Zemanick/Sweat Test")
source("~/GitHub/CF/get_genotype_class_severity.R")
```

```{r}
# Import exclusion data from Excel
diagnoses <- read_excel("./Data_Cleaned/AM reclassifications All Patients with CF or Non-CF.xlsx")
exclude <- diagnoses$PatientID[!is.na(diagnoses$Exclude)]
# Import data
df <- read.csv("./Data_Raw/Copy of Sweat Tests_AllFor202558_revised01Feb2021.csv", na.strings = "")
# Remove exclusions from Angela
df <- df[!(df$PatientID %in% exclude), ]
df$Diagnosis <- diagnoses$Diagnosis[match(df$PatientID, diagnoses$PatientID)]
# Filter per Maxie's code
df <- df %>%
  filter(
    Age.at.Test.in.Years < 50,
    !(Diagnosis == "Non-CF" & PatientID %in%
      c(22, 23, 24, 29, 31, 474, 483, 484, 485, 505, 510, 785, 791, 792, 793))
  ) %>%
  distinct(.) %>%
  filter(Sex != "", Demographics.Available == "TRUE") %>%
  mutate(Age.at.Test.in.Days = Age.at.Test.in.Weeks * 7)
# Add modulator info
cftr <- read.csv("./Data_Raw/Copy of Report CFTR Modulators.csv")
df$Modulator <- apply(df, 1, function(r) {
  mod <- cftr[cftr$Patient.ID == as.numeric(r["PatientID"]), ]
  if (any(duplicated(mod$Age.at.Start..years.))) {
    mod <- mod[-which(duplicated(mod$Age.at.Start..years.)), ]
  }
  mod <- mod[order(mod$Age.at.Start..years.), ]
  age <- as.numeric(r["Age.at.Test.in.Years"])
  as.character(cut(age, c(-Inf, mod$Age.at.Start..years., Inf),
    labels = c("None", mod$Modulator), right = F
  ))
})
# Genotype info
df$geno1 <- get_cf_genotype_class_severity(df$Genotypes1)
df$geno2 <- get_cf_genotype_class_severity(df$Genotypes2)
df$geno1[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno2[df$Diagnosis == "Non-CF"] <- "Not CF"
df$geno_risk <- paste(df$geno1, df$geno2, sep = "_")
df$geno_risk <- factor(df$geno_risk,
  levels = c(
    "Mild_Severe", "Mild_Unknown", "Not CF_Not CF",
    "Severe_Mild", "Mild_Mild",
    "Severe_Severe", "Severe_Unknown", "Unknown_Severe",
    "Unknown_Unknown"
  ),
  labels = c(
    "Low", "Low", "Not CF", "Low", "Low", "High", "Unknown",
    "Unknown", "Unknown"
  )
)
# Merge left and right values
df$Amount <- apply(df[, c("Amount", "Amount_Left", "Amount_Right")], 1, function(r) {
  amt <- suppressWarnings(as.numeric(as.character(r)))
  mean(amt, na.rm = T)
})
df$Cl <- apply(df[, c("Cl", "Cl_Left", "Cl_Right")], 1, function(r) {
  cl <- suppressWarnings(as.numeric(as.character(r)))
  mean(cl, na.rm = T)
})
df$Na <- apply(df[, c("Na", "Na_Left", "Na_Right")], 1, function(r) {
  na <- suppressWarnings(as.numeric(as.character(r)))
  mean(na, na.rm = T)
})
df$K <- apply(df[, c("K", "K_Left", "K_Right")], 1, function(r) {
  k <- suppressWarnings(as.numeric(as.character(r)))
  mean(k, na.rm = T)
})
# Sort
df <- df %>% arrange(PatientID, Age.at.Test.in.Weeks)
```

# Table 1: Patient Demographics and Sample Characteristics

For those with two sweat electrolyte measures (left and right), the average of the two was used.

```{r results='asis'}
t1_df <- df %>%
  filter(Age.at.Test.in.Days >= 2) %>%
  group_by(PatientID) %>%
  arrange(Age.at.Test.in.Weeks) %>%
  filter(row_number() == 1)
t1 <- tableby(Diagnosis ~ Sex + Age.at.Test.in.Years + geno_risk + Amount + Cl + Na + K,
  data = t1_df
)
summary(t1,
  labelTranslations =
    list(Age.at.Test.in.Years = "Age (Years)", geno_risk = "Genotype Severity")
)
```

# Chloride and Diagnosis

## Table

```{r}
df$Cl_group <- cut(df$Cl, c(0, 30, 60, Inf), right = F)
t <- table(df$Cl_group, df$Diagnosis)
t <- rownames_to_column(as.data.frame.matrix(t))
colnames(t)[1] <- "Chloride Range"
kable(t, row.names = F)
```

## Plot

```{r}
ggplot(df, aes(x = Diagnosis, y = Cl)) +
  geom_boxplot() +
  theme_bw()
```

```{r}
# Write only those with multiple values
mult <- unique(df$PatientID[which(duplicated(df$PatientID))])
multiples <- df[df$PatientID %in% mult, ]
write.csv(multiples, file = "./Data_Cleaned/multiple_values.csv", na = "", row.names = F)
```
